---
title: "Phyloseq"
author: "Sogomonyan Karina"
date: "27 05 2021"
output: html_document
---

# Handoff to phyloseq
```{r include=FALSE}
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
library(ggplot2); packageVersion("ggplot2")
library(dplyr)
library(vegan)
library(DESeq2)
library(gtools)
library(stringr)
library(tidyr)
theme_set(theme_bw())
```

```{r include=FALSE}
seqtab.nochim <- read.csv('Seq_nonchim')
```


```{r echo=TRUE}
samples.out <- rownames(seqtab.nochim)
subject <- sapply(strsplit(samples.out, "D"), `[`, 1)
samdf <- data.frame(Subject=subject)
rownames(samdf) <- samples.out
```

### Add information about an objective response in treatment
```{r echo=TRUE}
sam_d <- read.csv('/home/karina_mermaid/tools/1cloud/onko_metagenome/samp_dt.csv')
sam_d  <-  sam_d[c(2,3)]
sam_d <- sam_d %>%  add_row(filename_part1 = '75', Объективный.ответ = NA)

mer <- merge(samdf, sam_d, by.x="Subject", by.y = 'filename_part1')
mer <- mer %>% rename(Response = Объективный.ответ )
mer <- mer %>% mutate_if(is.character, list(~na_if(.,""))) 
rownames(mer) <-  samples.out
```


### Do phyloseq object
```{r echo=TRUE}
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(mer), 
               tax_table(taxa))

head(sample_data(ps)$Response, 10)
```


### Transform the table so as to get ASV numbers in the table with taxa, but the sequences themselves remained in the object
```{r echo=TRUE}
ps <- prune_samples(sample_names(ps), ps) # Remove mock sample
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps
```

### Rarecurve for samples
```{r}
# Shows how the variety changes depending on the size of the library (sample)
rarecurve((otu_table(ps)), step=100, cex=0.5)

```

Judging by the graph, the richness is highly dependent on the size of the library, so normalization is necessary


--------------------------------------------------------------------------------------------------------------------------
### Normalization with Deseq2

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Convert phyloseq object to deseq2 object 
ps_n <- subset_samples(ps, Response != "None")
ps_n <- prune_samples(sample_sums(ps_n) > 500, ps_n)
ps_n
diagdds = phyloseq_to_deseq2(ps_n, ~ Response)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
diagdds = estimateSizeFactors(diagdds, type="poscounts")
diagdds = estimateDispersions(diagdds, fitType = "local")
diagdds = DESeq(diagdds, fitType="local")
pst <- varianceStabilizingTransformation(diagdds) #This function calculates a variance stabilizing transformation (VST) from the fitted dispersion-mean relation(s) and then transforms the count data (normalized by division by the size factors or normalization factors), yielding a matrix of values which are now approximately homoskedastic.
```

### Assigning normalized values to the phyloseq object
```{r echo=TRUE, message=FALSE, warning=FALSE}
pst.dimmed <- t(as.matrix(assay(pst)))
pst.dimmed[pst.dimmed < 0.0] <- 0.0
ps.varstab <- ps_n
otu_table(ps.varstab) <- otu_table(pst.dimmed, taxa_are_rows = FALSE)
```

------------------------------------------------------------------------------------------------------------------------
### Built a barplot according to the top 20 plylum
```{r echo=TRUE}
top20 <- names(sort(taxa_sums(ps.varstab), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(ps.varstab, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)
plot_bar(ps.top20, x="Subject", fill="Phylum")
```

### Get the abundance table with taxanomic information of each ASV for each subject

```{r message=FALSE, warning=FALSE}
family<-as.data.frame(tax_table(ps.varstab))
short_tax<-family[1:5]
short_otu<-as.data.frame(otu_table(ps.varstab))
short_otu_ps <-as.data.frame(otu_table(ps.varstab))
short_tax_T <- as.data.frame(t(short_tax))
short_otu_T <- as.data.frame(t(short_otu))
info<-cbind(short_tax, short_otu_T)
res_phylum <- info %>% group_by(Phylum) %>% summarise_if(.predicate = is.numeric, funs(sum))
res_family <-info %>% group_by(Family) %>% summarise_if(.predicate = is.numeric, funs(sum))
res_order <-info %>% group_by(Order) %>% summarise_if(.predicate = is.numeric, funs(sum))
```


### Сonvert the number to percent
```{r echo=TRUE, message=FALSE, warning=FALSE}
data_percentage <- as.data.frame(apply(res_phylum[,2:66], 2, function(x){x*100/sum(x)}))
data_percentage_with_phylum <- cbind(res_phylum[, 1], data_percentage)

```

### Convert to long datatype and plot percentage stacked barplot
```{r}
# Convert into long data
long_data<-data_percentage_with_phylum %>%
   pivot_longer(!Phylum, names_to = "samples", values_to = "percent")
```



```{r}
sortered <- mixedsort(long_data$samples)
long_data_rearrenge <- long_data[mixedorder(long_data$samples),]

ggplot(long_data_rearrenge, aes(fill=Phylum, y=percent, x = samples)) + 
   geom_bar(stat="identity",  width=13) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_x_discrete(limits=long_data_rearrenge$samples)
```





















